Paper NA

### 2017-04-27 Paper_NA_br.tex
270417

	% Gung 2003 Positive anomalies of dlnxi down to 400km in the craton.
		\paragraph{age of the crust and its corresponding lithosphere}
		Our study confirms the presence of a thick lithospheric root in the craton with faster than average $V_S$ down to 250~km depth, with highest velocities, $\sim$4.7~km.s$^{-1}$, located at depths shallower than  150 km (e.g. Figure~\ref{cratoncrossedge} and \ref{cratoncrosscenter}). 
		Laterally, the cratonic root in $V_S$ coincides roughly with precambrian aged crust (see Figure~\ref{3d-VS} from 60 to 150~km deep and Figure~\ref{cratoncrossedge}.a ).
		The western edge of the craton follows the RMF (black thick dashed line in Figure~\ref{mapgeol}), instead of the western breakup line (See the red thin line in Figure~\ref{mapgeol} which coincides with RMF in Canada, but is diverging westward in the U.S.). 
		While the eastern/south-eastern coincides with the Llano-Grenville Front (See the blue thin line in Figure~\ref{mapgeol}). 
		The northern part ends at the coastline of the Arctic ocean including the arctic archipelago and Greenland.

		These lateral edges, coinciding with precambrian crust, are well observed down to 150 km depth. 
		Below that, the lithospheric core retracts toward the center of the precambrian continent comprising north western Juvenile arcs and orogens (e.g. Wopmay and BHT), north-eastern part of Juvenile crusts (e.g Yavapai/Mazatzal/Gr. Rhyolite), Slave, Hearne/Rae and western Superior cratons; while eastern Superior and Wyoming are cast out, as shown in Figure \ref{3d-VS}. 
		There are no obvious correlations between the age of a province and the amplitudes of anomalies and their extension in depth. 
		For example, archaean eastern Superior craton has $\sim 0\% \: V_S$ anomalies below 200 km (compared to $\sim+$ 4 $\%\:V_S$ beneath the western Superior), similar to the proterozoic Grenville province and archaean Wyoming ( also observed in Gao et al. 2002, Griffin et al. 1998, Menzies et al. 1993, Menzies et al. 2007); while the proterozoic belt of THO shows as high anomalies at the same depth as within the archaean Slave or western Superior lithosphere. 
		It is interesting to notice that at shallower depths than 100 km, the THO which welded the Superior and Rae/Hearne cratons \citep{hoffman1988united}, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding. 
		The THO comprises the archaean aged Sask craton (See Figure \ref{mapgeol}).
		The mantle beneath this exotic craton is interpreted to have been detached and replaced by both the Superior and the Hearne mantle lithosphere during final collision of the orogen \citep{nemeth2005lithospheric}. 
		This interpretation, according to \cite{faure2011seismic} can be extrapolated farther to the east and may explain why Trans-Hudson orogen has no imprint in the upper mantle.

		As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral negative variations $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans (See figure \ref{3d-Xi} where $dln(\xi)$ are shown according the Isotropic case ($\xi = 1$)).

		Below 250km, it is interesting to note that there is a large band of positive $dln(\xi)$ crossing the whole continent (across the US), making the connection between the Pacific and Atlantic mantles. 
		This feature is not present in most of Canada, where negative $dln(\xi)$ structure persist down to 400km and can be linked with the deep strong $\xi$ layer of \cite{gung2003global}. Our higher resolution model shows that this layer of strong $\xi$ may not be present beneath the whole cratonic lithosphere and it seems that this feature is marking a connection with Low Velocity Fingers (LVF) observed by \cite{french2013waveform}. % Below, this feature is still slightly visible beneath the US, while negative values of $dln\xi$ are observed beneath Canada
		% In particular, surface wave anisotropy studies show that cratons are underlain by a layer of radial anisotropy, where horizontally polarized shear waves are faster than vertically polarized shear waves in the depth range of 250–400 km 
		Generally, $\xi$ always appears higher than 1. 
		Except between 60 and 100 km, lower than 1 of $\xi$ ($V_{SV} > V_{SH}$) are observed in Mexico, at the southern part of Florida, the southern part of the Superior craton comprising the Mid Continental rifting, the southern part of the Granit Rhyiolite Province comprising the Failed Rifting area, a large area comprising southern Hearne, Medicine hat, northern Wyoming, B.H.T, Cordillera; and finally beneath Baffin Island. 
		% This feature can be observed in Figure \ref{3d-Xi} where perturbations are shown according the Isotropic case ($\xi = 1$)

		\paragraph{Lithospheric layering}
		Everywhere within the lithospheric core, from 60 to 100-150 km $V_S$ increases with depth (positive radial gradient) with a maximum of $\sim$4.7-4.8 km.s$^{-1}$ between 100 and 150 km (See Figure~\ref{cratoncrosscenter}). This can be an indication of a Mid lithospheric discontinuity within the lithopsheric core. 
		% However, MLD imaging is beyond the scope of this study.  
		Deeper, a negative gradient of $V_S$ is observed, marking the base of the lithospheric core \citep{eaton2009elusive}, with a minimum of $\sim$4.6 km.s$^{-1}$ reached at $\sim$270 km depth. 
		Such structure is very well illustrated at the center of the continent under the Slave, Hearne/Rae, THO and western Superior cratons; which is in contrast with the surrounding of the core, where velocities decrease with depths down to at most 100-150 km and deeper increase down to 400 km (see Figure~\ref{cratoncrosscenter} along the FF' and GG' cross sections of radial $V_S$ gradient showing that within the lithospheric core of NA is characterized by a positive gradient down to $\sim$150~km ,and then, a negative gradient down to at least $\sim$250~km.). 
		Beneath other Archean provinces like Wyoming craton (see Figure~\ref{cratoncrosscenter} along the FF' cross section), west to the RMF, such structure is not present. 
		Velocities decrease down to 120 km and then increase deeper than 400 km. 
		East of the RMF in Wyoming craton, velocities are monotically decreasing down to 250 km. 
		Also, beneath the eastern Superior craton, high velocities are confined at shallower depths than 100 km, and the radial structure is monotically decreasing with depth. 
		% The lithospheric core in North
		While the base of the lithosphere (and top of the low velocity zone - LVZ) shows strong topography, the bottom of the LVZ appears largely flat, in particular under the Slave, Hearne/Rae and western Superior cratons (See cross sections in Figure~\ref{cratoncrosscenter} and \ref{cratoncrossedge}.

		Regarding the radial evolution of $\xi$, cross section of the $\frac{d\xi}{dz}$ in Figure~\ref{cratoncrosscenter} shows alternances of trends.
		$\xi$ generally decrease from 50 to $\sim$70-100 km, then increase down to 150 km, decrease again down to $\sim$220-250 km and finally decrease to the Isotropy down to the Transition zone. Compared to $\frac{dV_S}{dz}$, $\frac{d\xi}{dz}$ in terms of amplitude is 4 order of magnitude below. It appears the band of negative $\frac{d\xi}{dz}$ between 150 and 250 km shows a topography at its upper limit. Around the lithopheric core, the band is $\sim$100 km thick between 150 and 250 km deep; while within the lithospheric core, the band is $\sim$50 km thick between 170 and 220 km deep. This feature might be a characteristic of the $\xi$ structure beneath a lithospheric core which needs to be confirmed in other cratonic studies.

@@

	\subsection{Basal traction beneath the lithosphere}
	% Its thickness is defined by the limit at which the cooler temperatures of the boundary layer can be resolved from the mantle adiabat (which is pinned to the potential temperature of the mantle, i.e., the temperature of the convecting mantle if it were adiabatically decompressed to the surface of Earth in the solid state).

@@

	Compare studies between ReOs and UPb and what they indicate
		\paragraph{Superior Craton}
			In our model, the Superior craton can be split in two parts: The western part showing typical cratonic structures while the eastern part shows slower velocities at all depths. Below 200 km, the distinction between the eastern and western part correlates the GMT.

		\paragraph{Western Ontario part}
			The western Superior craton shows typical lithospheric craton signatures with highest velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}. In terms of $\xi$, slight negative to zero perturbations are observed at the very western part, while close to the Mid continental rifting, highest negative perturbations indicate that $V_{SV}$ is higher than $V_{SH}$. This last feature is associated with the thinning of the cratonic lithosphere when crossing the Grenville front as observed by \cite{darbyshire2007new}.
			In the entire region there is a positive velocity gradient with a maximum of ca. $4.7 \: km.s^{-1}$ underneathed by a negative gradient  with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep.
% 
			% associated with a maximum of $\xi$ at ca. $150km$ deep.

			Opposed to the Slave craton, the topography of the top of the negative gradient is variable. 
			From SW to NE, the core of maximum velocity can reach beween 70 and 150km of thickness. 
			Beneath the maximum velocity region of  $\sim$ 150 km the negative gradient appears stipper than beneath the $70km$ one, while the topography of the deeper positive gradient appears flat. The topography of the bottom of the negative gradient is observed flat {\color{red} It is the same in Slave - One linkk to share with other regions}

			% The band of minimum velocity within the cratonic root has similar amplitudes, in terms of velocity, than beneath the younger lithosphere but at shalower depths between $100$ and $150 \: km$. 
			Following the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interesting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons. 
			Xenoliths analysis from kimberlites at Kirkaland Lake (south eastern part of the western Superior) exhibit some evidence of diamonds, i.e, the base of the lithosphere was deeper than 170 km at the time of kimberlitic eruption \citep{jones2014electrical} estimated at $160 - 152 \: Ma$ by \cite{heaman2000timing,heaman2003timing}. 
			This can be interpreted as either the lithosphere is today thick, and the lower lithosphere was not sampled by xenolith entrainment in the kimberlitic magmas (but nevertheless diamonds were entrained); or the lowermost lithosphere has been eroded away since eruption, possibly by the movement of the North American plate over the Great Meteor hotspot \citep{heaman2000timing,jones2014electrical}. In the study of \cite{faure2011seismic}, 23 $\%$ of kimberlite samples from southern Superior region contain diamonds. In contrast with 78 $\%$ at the northern Superior. This means that the diamond stability field within the mantle has been deteriorated through time and kimberlites that have passed across the mantle during/after such deterioration have less chances to intercept diamonds. \citep{faure2011seismic}

			\cite{kaban2015cratonic} proposed that basal drag induced by plate motion caused a south-western shift of the Superior's cratonic root. A feature that was observed by the local inversion of gravity, topography, crustal structure and seismic data.

			This basal deformation would have been enabled by the reheating of the lithosphere due to the passage of the Great Meteor hotpsot \citep[see][]{eaton2007seismic,kaban2015cratonic}.
			Our observations show that despite the archaean age of the Superior craton, the lithospheric root thickness of a craton is not flat and can be altered as broadly observed within the eastern part of the Superior craton.

@@

		\paragraph{Eastern Quebec part}
			Compared to its western part, the structure here is different than typical cratonic structure as shown in figure ~\ref{cratoncrossedge} 
			Fast velocities are present down to 150 km marking the shelf of the lithospheric nucleus. 
			Below, zero to slight positive anomalies caracterize the eastern superior craton (as in the wedded Grenville belt) showing the retractation toward the continent's center of lithospheric nucleus. A feature that was not present in our previous less well resolved models \citep{marone2007three,yuan20113,yuan2014multiscale}.
			Except at its south western part, negative gradient in terms of velocity is present from 50 to 250 km deep. 
			Whereas at the center of the continent, velocities are usually increasing (positive gradient) between 50 and 150 km, here it monotically decreases.
			The negative gradient is steeper and the minimum is radially thicker (more than 100 km thick compare to $\sim$50 km beneath the craton nucleus).

			Observed also by Villemaire 2012 with P-wave arrival time. Nice discussion in the end of the paper about lithospheric thinning.

@@

\subsubsection{Surrounding oceanic}

	Our inversion is focusing on the continental part, as shown in our spatial parameterization (i.e. see figure~\ref{density_coverage}) where the region is sampled by data.
	Except the oceanic borders of NA, further away, the structure has not been updated from SEMucb\_wm1 \citep{french2014whole}.

>>>>

	% Gung 2003 Positive anomalies of dlnxi down to 400km in the craton.
		\subsubsection{Lateral edges}
		Our study confirms the presence of a thick lithospheric root in the craton with faster than average $V_S$ down to 250~km depth, with highest velocities, $\sim$4.7~km.s$^{-1}$, located at depths shallower than  150 km (e.g. Figure~\ref{cratoncrossedge} and \ref{cratoncrosscenter}). 
		Laterally, the cratonic root in $V_S$ coincides roughly with precambrian aged crust (see Figure~\ref{3d-VS} from 60 to 150~km deep and Figure~\ref{cratoncrossedge}.a ).
		The western edge of the craton follows the RMF (black thick dashed line in Figure~\ref{3d-VS}), instead of the western breakup line (See the red dashed line in Figure~\ref{3d-VS} which coincides with RMF in Canada, but is diverging westward in the U.S.). 
		While the eastern/south-eastern coincides with the Llano-Grenville Front (See the blue thin line in Figure~\ref{3d-VS}) instead of the 650 Ma easter rift margin as observed in \cite{yuan2014multiscale} and \cite{schaeffer2014imaging}. 
		However at 60~km, fast velocities are in better agreement with the 650 Ma easter rift margin. 
		The northern part ends at the coastline of the Arctic ocean including the arctic archipelago and Greenland.

		\subsubsection{age of the crust and its corresponding lithosphere}
		These lateral edges, coinciding with precambrian crust, are well observed down to 150 km depth. 
		Below that, the lithospheric core retracts toward the center of the precambrian continent comprising north western Juvenile arcs and orogens (e.g. Wopmay and BHT), north-eastern part of Juvenile crusts (e.g Yavapai/Mazatzal/Gr. Rhyolite), Slave, Hearne/Rae and western Superior cratons; while eastern Superior and Wyoming are cast out, as shown in Figure \ref{3d-VS}. 
		This is consistent with the global study of \cite{bozdag2016global} but in contrast with recent regional \cite[e.g.][]{yuan2014lithospheric,yuan20113,schaeffer2014imaging,nettles2008radially} and global \cite[e.g][]{french2014whole,french200} studies, where the core is located at the north-east part of the continent.

		There are no obvious correlations between the age of a province with amplitude of anomalies and their extension in depth. Such observation is in agreement with several studies ( \cite[e.g for Baltic shield study][]{bruneton2004complex}, \cite[for Australian shield][]{fishwick2005contrasts},\cite[e.g for north American shield study][]{villemaire2012p,darbyshire2013seismic,porritt2015lithospheric} ).

		For example, in~\ref{3d-VS}, archaean eastern Superior craton has $\sim 0\% \: V_S$ anomalies below 200 km (compared to $\sim+$ 4 $\%\:V_S$ beneath the western Superior), similar to the proterozoic Grenville province as observed by \cite{villemaire2012p,bozdag2016global} and archaean Wyoming ( also observed in Gao et al. 2002, Griffin et al. 1998, Menzies et al. 1993, Menzies et al. 2007); while the proterozoic belt of THO at its western part, shows as high anomalies at the same depth as within the archaean Slave or western Superior lithosphere.

		Everywhere within the lithospheric core, from 60 to 100-150 km $V_S$ increases with depth (positive velocity gradient with depth) with a maximum of $\sim$4.7-4.8~km.s$^{-1}$ between 100 and 150 km (See Figure~\ref{cratoncrosscenter} and \ref{cratoncrossedge}).
		Deeper, a negative gradient of $V_S$ is observed, marking the base of the lithospheric core \citep{eaton2009elusive}, with a minimum of $\sim$4.6 km.s$^{-1}$ that can be reached down $\sim$270 km deep.
		We identify the cratonic root by the contour of $V_S$ = 4.68~km.s$^{-1}$ in Figure~\ref{cratoncrossedge}, which is chosen for clarity in order to show how variable the depth of the cratonic root is observed in our model.
		The $V_S$ = 4.68~km.s$^{-1}$ dashed line corresponds roughly to the middle of a zone of decreasing $V_S$ (negative velocity gradient with depth) which reaches a minimum almost uniformly around 250-270 km depth, although there are lateral variations in the strength of the minimum.

		\paragraph{Superior craton}
		The deepest roots are found in the Western and southwestern Superior craton where velocities in excess of 4.8~km.s$^{-1}$ are found at depths between 250 and 270 km (e.g. fig~\ref{cratoncrossedge} profiles BB' and EE'). 
		In contrast, the eastern Superior craton lithosphere is much thinner, around 150 km, with fastest velocities correspondingly peaking at depths less than 100 km as shown in fig~\ref{cratoncrossedge} profiles BB' and FF'. Such observations show that, the link between the age of the crust and the thickness of the lithospheric upper mantle beneath, is not straighforward. 
		% But implies a late alteration of the Superior craton since its cratonization in archaean times. See next part...

		\paragraph{Slave craton}
		In the north, the Slave cratonic lithosphere is $\sim$250~km thick (see Fig~\ref{cratoncrossedge}, cross sections AA' and DD'), but does not show as fast velocities as within the southwestern/western superior craton. 
		With shear velocities in the depth range 100-150 km not exceeding 4.7~km.s$^{-1}$. A value slower than in \cite{chen2007new}. I only have studies from RF so far(Snyder etc...), I am digging for more recent papers...

		\paragraph{THO}
		At the western part (Canadian part only) of the THO, the lithosphere is 250~km thick (see fig~\ref{cratoncrossedge} profiles AA' and FF' and profile CC' in fig.~\ref{cratoncrosscenter}), while a 100~km thinning is observed at the eastern part (i.e beneath the Hudson bay)

		It is interesting to notice that at shallower depths than 100 km, the western part of the THO which welded the Superior and Rae/Hearne cratons \citep{hoffman1988united}, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding. 
		The western part of the THO comprises the archaean aged Sask craton (See Figure \ref{mapgeol}).
		The mantle beneath this exotic craton is interpreted to have been detached and replaced by both the Superior and the Hearne mantle lithosphere during final collision of the orogen \citep{nemeth2005lithospheric}. 
		This interpretation, according to \cite{faure2011seismic} can be extrapolated farther to the east and may explain why Trans-Hudson orogen has no imprint in the upper mantle. However, the Hudson bay's local study from \cite{darbyshire2013seismic} , showed that the THO shows a channel of slow velocities compare to its surrounding down to 180~km deep.

		\paragraph{Yavapai/Mazatzal/Granit-Rhyolite orogens}
		The southernmost extension of the craton (Yavapai/Mazatzal/Granit-Rhyolite orogens) represent zones of accretion for Proterozoic material and mark the amalgamation of the archean cores \citep{hoffman1988united}. It shows a weaker but deep root (cross sections EE' and CC' in fig. ~\ref{cratoncrossedge} and map at 250 km depth in fig. \ref{3d-VS}) with velocities of almost 4.7~km.s$^{-1}$. There is some indication of incursion from rifting (section EE') separating these younger crustal regions from the Superior craton.	
		It is interesting to note that, beneath Nebraska and Kansas, the high velocity anomaly is located where cretaceous aged kimberlites are sampled. 
		These kimberlites indicate a lithosphere thickness of about 160~km and have been affected by asthenospheric upwelling related to the Mid-Continent Rift System. \citep{griffin2004lithosphere}

		The minimum in velocity is present everywhere at a roughy constant depth beneath the cratons while the base of this low velocity zone (LVZ) is found everywhere, including outside of the cratons between 300 and 325 km depth. That base is very smooth, compared to the structures above, and its minor undulations mirror the thickness of the lithosphere above, with slight upwarping beneath the deepest cratons.
		The LVZ is very thin and velocity reductions are small (min. ~4.65~km.s$^{-1}$) beneath regions where the lithosphere is thicker than 200 km, but thicker and more pronounced under the eastern superior (min. ~4.58~km.s$^{-1}$). In the regions bordering the cratons, i.e. outside of RMF and Grenville Front, the low velocity minimum is between 100-150~km and very pronounced ($V_S$ < 4.4~km.s$^{-1}$), and the transition out of it is more gradual, extending over ~150 km depth to reach velocities of 4.7~km.s$^{-1}$.

		\subsubsection{Radial anisotropy structure}
		As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral negative variations $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans (See figure \ref{3d-Xi} where $dln(\xi)$ are shown according the Isotropic case ($\xi = 1$)).

		Below 250km, it is interesting to note that there is a large band of positive $dln(\xi)$ crossing the whole continent (across the US), making the connection between the Pacific and Atlantic mantles. 
		This feature is not present in most of Canada, where negative $dln(\xi)$ structure persist down to 400km and can be linked with the deep strong $\xi$ layer of \cite{gung2003global}. Our higher resolution model shows that this layer of strong $\xi$ may not be present beneath the whole cratonic lithosphere and it seems that this feature is marking a connection with Low Velocity Fingers (LVF) observed by \cite{french2013waveform}. % Below, this feature is still slightly visible beneath the US, while negative values of $dln\xi$ are observed beneath Canada
		% In particular, surface wave anisotropy studies show that cratons are underlain by a layer of radial anisotropy, where horizontally polarized shear waves are faster than vertically polarized shear waves in the depth range of 250–400 km 
		Generally, $\xi$ always appears higher than 1. 
		Except between 60 and 100 km, lower than 1 of $\xi$ ($V_{SV} > V_{SH}$) are observed in Mexico, at the southern part of Florida, the southern part of the Superior craton comprising the Mid Continental rifting, the southern part of the Granit Rhyiolite Province comprising the Failed Rifting area, a large area comprising southern Hearne, Medicine hat, northern Wyoming, B.H.T, Cordillera; and finally beneath Baffin Island.

		Still need to do talk about the periodi upward of xi from depth to link with \ref{cratoncrossedge-xi}
		% This feature can be observed in Figure \ref{3d-Xi} where perturbations are shown according the Isotropic case ($\xi = 1$)

		% \paragraph{Lithospheric layering}
		% Everywhere within the lithospheric core, from 60 to 100-150 km $V_S$ increases with depth (positive radial gradient) with a maximum of $\sim$4.7-4.8 km.s$^{-1}$ between 100 and 150 km (See Figure~\ref{cratoncrosscenter}). This can be an indication of a Mid lithospheric discontinuity within the lithopsheric core. 
		% % However, MLD imaging is beyond the scope of this study.  
		% Deeper, a negative gradient of $V_S$ is observed, marking the base of the lithospheric core \citep{eaton2009elusive}, with a minimum of $\sim$4.6 km.s$^{-1}$ reached at $\sim$270 km depth. 
		% Such structure is very well illustrated at the center of the continent under the Slave, Hearne/Rae, THO and western Superior cratons; which is in contrast with the surrounding of the core, where velocities decrease with depths down to at most 100-150 km and deeper increase down to 400 km (see Figure~\ref{cratoncrosscenter} along the FF' and GG' cross sections of radial $V_S$ gradient showing that within the lithospheric core of NA is characterized by a positive gradient down to $\sim$150~km ,and then, a negative gradient down to at least $\sim$250~km.). 
		% Beneath other Archean provinces like Wyoming craton (see Figure~\ref{cratoncrosscenter} along the FF' cross section), west to the RMF, such structure is not present. 
		% Velocities decrease down to 120 km and then increase deeper than 400 km. 
		% East of the RMF in Wyoming craton, velocities are monotically decreasing down to 250 km. 
		% Also, beneath the eastern Superior craton, high velocities are confined at shallower depths than 100 km, and the radial structure is monotically decreasing with depth. 
		% % The lithospheric core in North
		% While the base of the lithosphere (and top of the low velocity zone - LVZ) shows strong topography, the bottom of the LVZ appears largely flat, in particular under the Slave, Hearne/Rae and western Superior cratons (See cross sections in Figure~\ref{cratoncrosscenter} and \ref{cratoncrossedge}.

		% Regarding the radial evolution of $\xi$, cross section of the $\frac{d\xi}{dz}$ in Figure~\ref{cratoncrosscenter} shows alternances of trends.
		% $\xi$ generally decrease from 50 to $\sim$70-100 km, then increase down to 150 km, decrease again down to $\sim$220-250 km and finally decrease to the Isotropy down to the Transition zone. Compared to $\frac{dV_S}{dz}$, $\frac{d\xi}{dz}$ in terms of amplitude is 4 order of magnitude below. It appears the band of negative $\frac{d\xi}{dz}$ between 150 and 250 km shows a topography at its upper limit. Around the lithopheric core, the band is $\sim$100 km thick between 150 and 250 km deep; while within the lithospheric core, the band is $\sim$50 km thick between 170 and 220 km deep. This feature might be a characteristic of the $\xi$ structure beneath a lithospheric core which needs to be confirmed in other cratonic studies.

@@

	\subsection{Basal traction beneath the lithosphere}

	Ici faire le lien avec les circle figure, fingers de Scott. Probleme pratique: La carte geol ne va pas jusquau MAR et donc afficher le track n'ira pas jusqau bout...  Avec la figure \ref{Vs-circle}

		As in \cite{schaeffer2014imaging,yuan2014lithospheric,james2014rayleigh,silveira2002anisotropic,mocquet1990three}, we observe along the negative anomaly structure following the Appalachian continental shelf, a stretched positive anomaly structure from the Suwanne terrane in Florida to the Avalonian terranes of New-Scotland. 
		At 60~km, it is made of several blocks of at most $+3\%$ anomalies emdedded in a $+2\%$ structure, associated with zero to positive $\xi$ anomalies. Below 100~km the pan african lithosphere \citep{yuan2014lithospheric} is cut in two by a slow velocity anomaly comprising both the Great Meteor Track (GMT) and Bermuda tracks, and further southeast, the Low Velocities Fingers (LVF) observed in \cite{french2013waveform}. This is best observed in figure~\ref{Vs-circle}. 
		Further East, the slow anomaly follows the GMT crossing the continental shelf, Appalachian and Grenville orogens down to 200~km. This slow anomaly corresponds to the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. We observe at 250~km that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.

		Our model images a slow velocity anomaly in agreement with the reconstruction of the Great Meteor Track proposed by \cite{heaman2000timing}.
		In figure \ref{cratoncrosscenter} the cross-section DD' follows the Great Meteor track and shows the thinning of the cratonic lithosphere that may be attributed to the passage of the Great Meteor hotspot during the opening of the atlantic ocean.
		It is interesting to note that the low velocity channel beneath the base of the lithosphere shows lower velocities than usualy observed beneath the other Archaean cratons.

		% \cite{heaman2000timing}, based on $U-Pb$ perovskite age determinations of kimberlites, proposed a progressive southeastward younging of kimberlite magmatism throughout much of eastern North America. Considering it as a strong support for small volume mantle melting that have occurred along the continental portion of the Mesozoic Great Meteor mantle plume hotspot track initiated during the opening of the North Atlantic Ocean at about $200Ma$. \citep[See figure 4. in][]{heaman2000timing}

		% In figure \ref{cratoncrosscenter} the cross-section DD' follows the Great Meteor track and shows the thinning of the cratonic lithosphere that may be attributed to the passage of the Great Meteor hotspot during the opening of the atlantic ocean.

		After accretion of the Appalachian terranes to eastern Laurentia during the Paleozoic, no further significant episodes of magmatism or tectonism affected the western Superior region until the passage of the Great Meteor hot spot at 200–110~Ma \cite{heaman2000timing,villemaire2012p}.
		Xenoliths analysis from kimberlites at Kirkaland Lake (south eastern part of the western Superior) exhibit some evidence of diamonds, i.e, the base of the lithosphere was deeper than 170~km at the time of kimberlitic eruption \citep{jones2014electrical} estimated at $160 - 152 \: Ma$ by \cite{heaman2000timing,heaman2003timing}. 
		This can be interpreted as either the lithosphere is today thick, and the lower lithosphere was not sampled by xenolith entrainment in the kimberlitic magmas (but nevertheless diamonds were entrained); or the lowermost lithosphere has been eroded away since eruption, possibly by the movement of the North American plate over the Great Meteor hotspot \citep{heaman2000timing,villemaire2012p,jones2014electrical}. In the study of \cite{faure2011seismic}, 23 $\%$ of kimberlite samples from southern Superior region contain diamonds. In contrast with 78 $\%$ at the northern Superior. This means that the diamond stability field within the mantle has been deteriorated through time and kimberlites that have passed across the mantle during/after such deterioration have less chances to intercept diamonds. \citep{faure2011seismic}

		\cite{davies1994thermomechanical} described a process of thermo-mechanical erosion which involves heating-softening of the lithospheric base by a plume tail, followed by mechanical removal of the material due to convective transfer. \citep{rondenay2000lithospheric}
		\cite{kaban2015cratonic} proposed that basal drag induced by plate motion caused a south-western shift of the Superior's cratonic root. A feature that was observed by the local inversion of gravity, topography, crustal structure and seismic data. This basal deformation would have been enabled by the reheating of the lithosphere due to the passage of the Great Meteor hotpsot \citep[see][]{eaton2007seismic,kaban2015cratonic}.

		Our observations show that despite the archaean age of the Superior craton, the lithospheric root thickness of a craton is not flat and can be altered as broadly observed within the eastern part of the Superior craton.

	% The largest continental keel on Earth is that beneath Laurentia, the cratonic core of North America, where the geological record spans more than 2 billion years of the Precambrian from the Archean formation of the Superior craton to the Proterozoic Grenville orogen [Hoffman, 1988]. After accretion of the Appalachian terranes to eastern Laurentia during the Paleozoic, no further significant epi- sodes of magmatism or tectonism affected the region until the passage of the Great Meteor hot spot at $200–110 Ma [e.g., Heaman and Kjarsgaard, 2000]. The hot spot track extends $6000 km from northern Hudson Bay to the northeast US coast, and continues southeast as a chain of Atlantic seamounts. Eastern North America thus represents an excellent opportunity to address questions associated with Precambrian tectonics, and the ability of the litho- sphere to resist modification by hot spot tectonism.

	% Its thickness is defined by the limit at which the cooler temperatures of the boundary layer can be resolved from the mantle adiabat (which is pinned to the potential temperature of the mantle, i.e., the temperature of the convecting mantle if it were adiabatically decompressed to the surface of Earth in the solid state).

@@

	% Is the plate motion slower beneath cratons, if yes maybe basal drag occurs there.

% 	Compare studies between ReOs and UPb and what they indicate
% 		\paragraph{Superior Craton}
% 			In our model, the Superior craton can be split in two parts: The western part showing typical cratonic structures while the eastern part shows slower velocities at all depths. Below 200 km, the distinction between the eastern and western part correlates the GMT.

% 		\paragraph{Western Ontario part}
% 			The western Superior craton shows typical lithospheric craton signatures with highest velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}. In terms of $\xi$, slight negative to zero perturbations are observed at the very western part, while close to the Mid continental rifting, highest negative perturbations indicate that $V_{SV}$ is higher than $V_{SH}$. This last feature is associated with the thinning of the cratonic lithosphere when crossing the Grenville front as observed by \cite{darbyshire2007new}.
% 			In the entire region there is a positive velocity gradient with a maximum of ca. $4.7 \: km.s^{-1}$ underneathed by a negative gradient  with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep.
% % 
% 			% associated with a maximum of $\xi$ at ca. $150km$ deep.

% 			Opposed to the Slave craton, the topography of the top of the negative gradient is variable. 
% 			From SW to NE, the core of maximum velocity can reach beween 70 and 150km of thickness. 
% 			Beneath the maximum velocity region of  $\sim$ 150 km the negative gradient appears stipper than beneath the $70km$ one, while the topography of the deeper positive gradient appears flat. The topography of the bottom of the negative gradient is observed flat {\color{red} It is the same in Slave - One linkk to share with other regions}

% 			% The band of minimum velocity within the cratonic root has similar amplitudes, in terms of velocity, than beneath the younger lithosphere but at shalower depths between $100$ and $150 \: km$. 
% 			Following the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interesting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons. 
% 			Xenoliths analysis from kimberlites at Kirkaland Lake (south eastern part of the western Superior) exhibit some evidence of diamonds, i.e, the base of the lithosphere was deeper than 170 km at the time of kimberlitic eruption \citep{jones2014electrical} estimated at $160 - 152 \: Ma$ by \cite{heaman2000timing,heaman2003timing}. 
% 			This can be interpreted as either the lithosphere is today thick, and the lower lithosphere was not sampled by xenolith entrainment in the kimberlitic magmas (but nevertheless diamonds were entrained); or the lowermost lithosphere has been eroded away since eruption, possibly by the movement of the North American plate over the Great Meteor hotspot \citep{heaman2000timing,jones2014electrical}. In the study of \cite{faure2011seismic}, 23 $\%$ of kimberlite samples from southern Superior region contain diamonds. In contrast with 78 $\%$ at the northern Superior. This means that the diamond stability field within the mantle has been deteriorated through time and kimberlites that have passed across the mantle during/after such deterioration have less chances to intercept diamonds. \citep{faure2011seismic}

% 			\cite{kaban2015cratonic} proposed that basal drag induced by plate motion caused a south-western shift of the Superior's cratonic root. A feature that was observed by the local inversion of gravity, topography, crustal structure and seismic data.

% 			This basal deformation would have been enabled by the reheating of the lithosphere due to the passage of the Great Meteor hotpsot \citep[see][]{eaton2007seismic,kaban2015cratonic}.
% 			Our observations show that despite the archaean age of the Superior craton, the lithospheric root thickness of a craton is not flat and can be altered as broadly observed within the eastern part of the Superior craton.

@@

		% \paragraph{Eastern Quebec part}
		% 	Compared to its western part, the structure here is different than typical cratonic structure as shown in figure ~\ref{cratoncrossedge} 
		% 	Fast velocities are present down to 150 km marking the shelf of the lithospheric nucleus. 
		% 	Below, zero to slight positive anomalies caracterize the eastern superior craton (as in the wedded Grenville belt) showing the retractation toward the continent's center of lithospheric nucleus. A feature that was not present in our previous less well resolved models \citep{marone2007three,yuan20113,yuan2014multiscale}.
		% 	Except at its south western part, negative gradient in terms of velocity is present from 50 to 250 km deep. 
		% 	Whereas at the center of the continent, velocities are usually increasing (positive gradient) between 50 and 150 km, here it monotically decreases.
		% 	The negative gradient is steeper and the minimum is radially thicker (more than 100 km thick compare to $\sim$50 km beneath the craton nucleus).

		% 	Observed also by Villemaire 2012 with P-wave arrival time. Nice discussion in the end of the paper about lithospheric thinning.

@@

% \subsubsection{Surrounding oceanic}

% 	Our inversion is focusing on the continental part, as shown in our spatial parameterization (i.e. see figure~\ref{density_coverage}) where the region is sampled by data.
% 	Except the oceanic borders of NA, further away, the structure has not been updated from SEMucb\_wm1 \citep{french2014whole}.
